Comparison of MCMC algorithms in Stochastic Volatility Models using Simulation Based Calibration

### 2023-10-10 manuscript/final_presentation/presentation.qmd
updated images

<!--Benefit of Bayesian model is that they're generative. Model can simulate data based on known parameter values.-->
<!-- Pick a value for parameteres, simulate the data, estimate the model on simulated data. See if you can capture true parameters -->
<!-- Phi not good! The model/algorithm doesn't estimate phi very well. -->
<!-- Not true. 5% chance that any random draw from the true probabiltiy dist will fall outside the 95% interval.-->

@@

## Kim, Shephard, Chib (1998)

$$
\begin{aligned}
y_t^{*} &= log(y_t^2) \\ 
&= log((\epsilon_t exp(h_t/2))^2) \\
&=  log(exp(h_t)) + log(\epsilon_t^2) \\
&= h_t + log(\epsilon_t^2)  \\
&= h_t + z_t \\
\end{aligned}
$$

Where $z_t = log(\epsilon_t^2)$ follows a log chi-squared distribution with mean -1.2704 and variance 4.93.

- Gaussian Mixture model
- Kalman Filter
- Conjugate posteriors
- Metropolis Hastings

## Hamiltonian Monte Carlo
- The Stan programming language's implementation of Hamiltonian Monte Carlo, the No-U-Turn Sampler (Hoffman & Gelman 2014), will be compared with KSC's strategy.

- Key innovations:

    + Uses the gradients of the target posterior distribution to generate an efficient path for the sampler to explore.

    + Allows for direct sampling of the specified stochastic volatility model

    + Own programming language. Flexibly estimate any model

@@

## HMC (1000 Iterations)

## HMC (5000 Iterations)

## KSC (1000 Iterations)

## KSC (5000 Iterations)

@@

## Reparameterised HMC (5000 Iterations)

## Reparameterised KSC (5000 Iterations)

# Discussion

# Conclusion

>>>>

<!--Benefit of Bayesian model is that they're generative. Model can simulate data based on known parameter values.-->
<!-- Pick a value for parameteres, simulate the data, estimate the model on simulated data. See if you can capture true parameters -->
<!-- Phi not good! The model/algorithm doesn't estimate phi very well. -->
<!-- Not true. 5% chance that any random draw from the true probabiltiy dist will fall outside the 95% interval.-->
<!-- Gen models are great because we can simulate data from the process implied by the model conditional on parameters. -->

@@

## Markov Chain Monte Carlo (MCMC) algorithms

**Kim, Shephard, Chib (1998)**

$$
\begin{aligned}
y_t &= \space e^{h_t/2} \epsilon_t \\
y_t^{*} &= log(y_t^2) = h_t + z_t\\ 
\end{aligned}
$$

- Approximate log chi squared error using Gaussian mixture model
- Kalman Filter
- Conjugate posteriors
- Metropolis Hastings

**Hamiltonian Monte Carlo**

- Own programming language. Flexibly estimate any model

- Uses the gradients of the target posterior distribution to generate an efficient path for the sampler to explore.

- Allows for direct sampling of the specified stochastic volatility model

@@

## Results
- 5000 SBC iterations
- Simulated datasets of 1000 (which means 1000 latent states)
- 999 posterior draws for HMC (1000 rank statistics per parameter)
- 9,999 posterior draws for KSC (10,000 rank statistics per parameter)

## HMC (5000 Iterations)
![]("../results/hmc_cp_5k.png")

## KSC (5000 Iterations)
![]("../results/ksc_cp_5k.png")

@@

## Reparameterised HMC (5000 Iterations)
![]("../results/hmc_ncp_5k.png")

## Reparameterised KSC (5000 Iterations)
![]("../results/ksc_ncp_5k.png")

## Interpretation of miscalibration

## Discussion & Limitations
- Reparameterised model with HMC performs best based on calibration

- SBC is computationally intensive (required use of Monash HPC Cluster)

- What constitutes a fair comparison? Reparameterised HMC calibrated after 999 draws.

# Questions
